---
title: "Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
```

Import that data and turn into factors,

```{r}
df <- read_csv("../data/clean_survey_data.csv")
fac_df <- df %>% 
  mutate(first_lang = as.factor(first_lang),
         first_lang_time = as.factor(first_lang_time),
         first_lang_method = as.factor(first_lang_method),
         preference = as.factor(preference),
         academic_background = as.factor(academic_background))

# remove the time component
dat <- fac_df %>% 
  select(-time)
```

Get contingency table of first language and preferred language,

```{r}
# get counts of combined first language/preferred language pairs
# note some combinations don't have a representative and so aren't in the table
lang_pref_counts <- dat %>% 
  group_by(first_lang, preference) %>% 
  summarise(count=n())

v <- lang_pref_counts$count

# add 0 entries for missing pairs
entries <- c(v[1:2], 0, v[3:8], 0, v[9:13], 0, v[14:15], 0, v[16:17])

# turn list of entries into contingency table
cont_tbl <- matrix(entries, ncol=3, byrow=TRUE)
colnames(cont_tbl) <- unique(lang_pref_counts$preference)
rownames(cont_tbl) <- unique(lang_pref_counts$first_lang)

cont_tbl

# write  to file
save(cont_tbl, file="../results/dat/cont_tbl.rda")
```


Perform exact test,

```{r}
fisher.test(cont_tbl) %>% broom::tidy()
```

Encode data in terms of preferences,

```{r}
# dat with col for preference/non-preference (preference can be R OR Python)
dat_has_pref <- dat %>% 
  mutate(has_pref = ifelse(preference %in% c("R", "Python"), 1, 0),
         has_pref = as.factor(has_pref))

# dat with col for R preference -- 0 corresponds to Python preference or no preference
dat_pref_R <- dat %>% 
  mutate(pref_R = ifelse(preference == "R", 1, 0),
         pref_R = as.factor(pref_R))

# dat with col for Python preference -- 0 corresponds to R preference or no preference
dat_pref_Py <- dat %>% 
  mutate(pref_Py = ifelse(preference == "Python", 1, 0),
         pref_Py = as.factor(pref_Py))
```

Running logistic regressions,

**Having Preference Related to First Language**

```{r warning=FALSE}
dat_has_pref$first_lang <- relevel(dat_has_pref$first_lang, ref="R")

# with  all confounders
glm(has_pref ~ . - preference, family = binomial, data=dat_has_pref) %>% broom::tidy()

glm(has_pref ~ . - preference - academic_background, family = binomial, data=dat_has_pref) %>% broom::tidy()

glm(has_pref ~ . - preference - academic_background - first_lang_time, family = binomial, data=dat_has_pref) %>% broom::tidy()

glm(has_pref ~ . - preference - academic_background - first_lang_time - first_lang_method, family = binomial, data=dat_has_pref) %>% broom::tidy()

# no counfounders
glm(has_pref ~ first_lang, family = binomial, data=dat_has_pref) %>% broom::tidy()
```

Try releveling `first_lang`,

```{r warning=FALSE}
dat_has_pref$first_lang <- relevel(dat_has_pref$first_lang, ref="Python")

# with  all confounders
glm(has_pref ~ . - preference, family = binomial, data=dat_has_pref) %>% broom::tidy()

glm(has_pref ~ . - preference - academic_background, family = binomial, data=dat_has_pref) %>% broom::tidy()

glm(has_pref ~ . - preference - academic_background - first_lang_time, family = binomial, data=dat_has_pref) %>% broom::tidy()

glm(has_pref ~ . - preference - academic_background - first_lang_time - first_lang_method, family = binomial, data=dat_has_pref) %>% broom::tidy()

# no counfounders
glm(has_pref ~ first_lang, family = binomial, data=dat_has_pref) %>% broom::tidy()
```


Even as we remove confounders that appear to be highly insignificant, we see only modest decreases in the p-values of other variables. Ultimately, even just regressing on `first_lang` doesn't have any significant coefficients for either of the ways that we releveled it.


**R Preference Related to First Language**

```{r warning=FALSE}
dat_pref_R$first_lang <- relevel(dat_pref_R$first_lang, ref="Python")

# with confounders
glm(pref_R ~ . -preference, family=binomial, data=dat_pref_R) %>% broom::tidy()

# removing first lang learning method
glm(pref_R ~ . -preference - first_lang_method, family=binomial, data=dat_pref_R) %>% broom::tidy()

# remove time from learning first language
glm(pref_R ~ . -preference - first_lang_method - first_lang_time, family=binomial, data=dat_pref_R) %>% broom::tidy()

# remove academic background
glm(pref_R ~ . -preference - first_lang_method - first_lang_time - academic_background, family=binomial, data=dat_pref_R) %>% broom::tidy()

# no confounders
glm(pref_R ~ first_lang, family=binomial, data=dat_pref_R) %>% broom::tidy()
```

Releveling,

```{r warning=FALSE}
dat_pref_R$first_lang <- relevel(dat_pref_R$first_lang, ref="Other")

# with confounders
glm(pref_R ~ . -preference, family=binomial, data=dat_pref_R) %>% broom::tidy()

# removing first lang learning method
glm(pref_R ~ . -preference - first_lang_method, family=binomial, data=dat_pref_R) %>% broom::tidy()

# remove time from learning first language
glm(pref_R ~ . -preference - first_lang_method - first_lang_time, family=binomial, data=dat_pref_R) %>% broom::tidy()

# remove academic background
glm(pref_R ~ . -preference - first_lang_method - first_lang_time - academic_background, family=binomial, data=dat_pref_R) %>% broom::tidy()

# no confounders
glm(pref_R ~ first_lang, family=binomial, data=dat_pref_R) %>% broom::tidy()
```

**Python Preference Related to First Language**

```{r}
dat_pref_Py$first_lang <- relevel(dat_pref_R$first_lang, ref="Other")

# all confounders
glm(pref_Py ~ . - preference, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove time since first language learned
glm(pref_Py ~ . - preference - first_lang_time, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove academic background
glm(pref_Py ~ . - preference - first_lang_time - academic_background, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove method used to learn first language
glm(pref_Py ~ . - preference - first_lang_time - academic_background - first_lang_method, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove employment importance
glm(pref_Py ~ first_lang, family=binomial, data=dat_pref_Py) %>% broom::tidy()
```

Releveling,

```{r}
dat_pref_Py$first_lang <- relevel(dat_pref_R$first_lang, ref="R")

# all confounders
glm(pref_Py ~ . - preference, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove time since first language learned
glm(pref_Py ~ . - preference - first_lang_time, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove academic background
glm(pref_Py ~ . - preference - first_lang_time - academic_background, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove method used to learn first language
glm(pref_Py ~ . - preference - first_lang_time - academic_background - first_lang_method, family=binomial, data=dat_pref_Py) %>% broom::tidy()

# remove employment importance
glm(pref_Py ~ first_lang, family=binomial, data=dat_pref_Py) %>% broom::tidy()

py_no_conf <- glm(pref_Py ~ first_lang, family=binomial, data=dat_pref_Py) %>% broom::tidy()
save(py_no_conf, file="../results/dat/py_no_conf.rda")
```


